---
title: 'BP40087: Tumour IHC Analyses update MODIFIED plots'
author: "Eveline Nüesch- Mariana Bustamante"
date: "Data Snapshot: vad latest data 13.06.22"
output:
  html_document:
    theme: flatly
    code_folding : hide
    number_sections: false
    toc: true
    toc_float: true
    fig_width: 8
    fig_height: 6
    dpi: 300
    # fig_height: 9
    # toc_depth: 2
    # toc_float: 
    #   collapsed: false
    #   smooth_scroll: false
    #theme: spacelab
editor_options: 
  chunk_output_type: console
# params:
#   palette:
#     label: "Select color palette:"
#     value: dosecols_viridis
#     input: select
#     choices: [dosecols_viridis, dosecols, dosecols_quali]
  
---


```{r, include=FALSE}
##Empty the working memory
rm(list = ls())
#knitr::opts_chunk$set(echo=TRUE)
library(knitr)
library(tidyverse)
library(dplyr)
library(cowplot)
library(rice)
library(LaplacesDemon)
library(ggalluvial)
library(ggpubr)
library(survival)
library(survminer)




snapshot.date <- "13 June 2022"

set.seed(42)

dosecols <- c( "45" = "#ffffd9",
                   "90" = "#edf8b1",
                  "130" = "#c7e9b4",
                  "260" = "#7fcdbb",
                  "500" = "#41b6c4",
                  "670" = "#1d91c0",
                 "1000" = "#225ea8",
                 "2000" = "#0c2c84")

dosecols_quali <- c( "45" = "#ffff33",
                   "90" = "#a65628",
                  "130" = "#984ea3",
                  "260" = "#4daf4a",
                  "500" = "#377eb8",
                  "670" = "#f781bf",
                 "1000" = "#ff7f00",
                 "2000" = "#e41a1c")

palette_group <- c("RO7122290 single agent"= "#009E73", 
                   "RO7122290 + atezolizumab" = "#D55E00")

group_shape <- c("RO7122290 single agent"= 23, 
                   "RO7122290 + atezolizumab" = 21)


respcolours <-   c("CR" = "#009E73",
                    "PR"= "#0072B2", 
                           "SD" = "#F0E442", 
                           "PD" = "#D55E00"#, 
                    # "Missing" = "grey",
                    # "CR (Complete response)" = "#009E73", 
                    # "PR (Partial response)"= "#0072B2", 
                    # "PD (Progressive disease)"= "#D55E00",
                    # "SD (Stable disease)" = "#F0E442"
                   )

dosecols_viridis <- c(
  "45", "90", "130", "260", "500", "670", "1000", "2000"
) %>%
  set_names(viridis::viridis(length(.), direction = -1), .)


order_bor <- c("CR",
               "PR",
               "SD", 
               "PD", 
              #"NA", 
              "Missing")

              imm_pty_mapping <- c("EXCLUDED - INTRA-TUMORAL" = "Excluded", 
                   "DESERT" = "Desert",
                   "INFLAMED" = "Inflamed")

visit_mapping <- c("801_BASELINE" = "Baseline", 
                   "801_" = "Baseline",
                   "802_ON TREATMENT" = "On-treatment", 
                   "802_ON_TREATMENT" = "On-treatment", 
                   "800_ARCHIVAL_TISSUE" = "Archival",
                   "800_BASELINE" = "Archival",
                   "Tissue Sample Collection-Screening" = "Screening", 
                   "Tissue Coll. On-Treatment" = "On-treatment",
                   "801" = "Baseline", 
                   "802" = "On-treatment", 
                   "800" = "Archival"
                   )

                   
cols_allu <- c("CR" = "#009E73",
                           "SD" = "#F0E442", 
                           "PR"= "#0072B2", 
                           "PD" = "#D55E00", 
              "Missing" = "#d9d9d9",
              "Desert" = "#8dd3c7", 
              "Inflamed" = "#bebada",
              "Excluded" = "#ffffb3",
              "Fail" = "#F0E442", 
              "RO7122290 single agent" = "#e78ac3", 
              "RO7122290 + atezolizumab" = "#ffd92f" 
            )



# Reading external data
source("R/read_data_internal_mb.R")

```

```{r functions, echo=FALSE}


plot_ABS <- function(dat, title, y_label = "Density (cells/mm2)", palette = dosecols_viridis) {

    
    gg <- 
      ggplot(dat[dat$VISITNUM<=2,], aes(x=VISIT, y=AVAL)) +
      geom_boxplot(aes(group=VISIT), colour="gray14", fill="#f0f0f0", outlier.shape=NA, size=0.4) +
      geom_line(aes(x = VISIT,  group=USUBJID), alpha=0.25, size=0.5, position = position_dodge(0.3)) +
      geom_point(aes(x=VISIT, fill = as.factor(EXDOSE),  group=USUBJID), size=2.5, alpha=0.75, shape=21, col = "black", position = position_dodge(0.3))+
      
      scale_y_log10() +
      xlab(" ") +
      ylab(y_label) +
      scale_fill_manual(name="RO7122290 dose (mg)", values = palette) +
      labs(title = title)  + 
      facet_grid(col = vars(GROUP)) +
      theme_bw()
    
    print(gg)
    
}
  
plot_CHB <- function(dat, title, 
                     y_label = "Fold change from baseline",
                     palette = dosecols_viridis){

    
    dat1 <- dat %>% filter(VISIT=='On Treatment' & is.na(RCHG)==FALSE)
  
    gg <- 
      ggplot(dat1, aes(x=GROUP, y=RCHG)) +
      geom_boxplot(aes(group=GROUP), colour="darkgray", fill="#f0f0f0", outlier.shape=NA, size=0.4) +
      geom_point(aes(x=GROUP, fill = as.factor(EXDOSE)), size=2.5, col = "black", alpha=0.75, shape=21, position = position_dodge(0.3)) +
      scale_y_log10() +
      xlab(" ") +
      ylab(y_label) +
      scale_fill_manual(name="RO7122290 dose (mg)", values = palette) +
      geom_hline(aes(yintercept = 2), linetype="dotted" , color = "black") +
      ggtitle(title) +
      theme_bw() 

    
    print(gg)
} 


table_ZB <- function(dat, title){
  
    dat.table = dat %>% filter(VISIT=='On Treatment') %>%
                mutate(EXDOSE = paste(as.character(EXDOSE), EXDOSU)) %>%
                mutate(BASE = paste(as.character(round(BASE)), PARAMU)) %>%
                mutate(AVAL =  paste(as.character(round(AVAL)), PARAMU)) %>%
                mutate(RCHG = round(RCHG, digits=2)) %>%
                select(USUBJID, MHTERM, EXDOSE, AVALC, BASE, AVAL, RCHG) 
    
    tt <- kable(dat.table,
          col.names = c("Subject",
                        "Tumour type",
                        "RO7122290 dose (mg)",
                        "Best Overall Response",
                        "Screening Value",
                        "On-Treatment Value",
                        "Fold induction"),
          caption = title)
    print(tt)
}
```

```{r echo=FALSE, warning=FALSE}
# CD8+ T cells infiltration (IRIS-IHC)
CD8T <- derive_ZB(sdtm$zb, '(MKI67+ CD8A+)+(MKI67- CD8A+)', 'DNPCTAR')
CD8T <- derive_pchgdata(CD8T)
# CD8+ T cells proliferation (IRIS-IHC)
KI67CD8T <- derive_ZB(sdtm$zb, 'MKI67+ CD8A+', 'DNPCTAR')
KI67CD8T <- derive_pchgdata(KI67CD8T)

FAP_pos <- derive_ZB(sdtm$zb, "FAP+", "STPOSARP") # this is in percentage
FAP_pos <- derive_pchgdata(FAP_pos)


FAP_ker_neg <- derive_ZB(sdtm$zb, "FAP+/KRT-", "STPOSARP") # this is in percentage
FAP_ker_neg <- derive_pchgdata(FAP_ker_neg) %>% 
                mutate(original_indication = str_replace(MHTERM, 
                                                      "LOCAL |METASTATIC ", 
                                            ""
                                              ), 
                   indication = case_when(original_indication == "NON SMALL CELL LUNG CANCER" ~
                                            "NSCLC", 
                                          original_indication == "SMALL CELL LUNG CANCER" ~
                                            "SCLC", 
                                          original_indication %in% c("RECTAL CANCER", 
                                                                     "COLON CANCER") ~
                                            "CRC", 
                                          original_indication %in% c("TONGUE CANCER", 
                                                                     "SALIVARY GLAND CANCER") ~
                                            "HNC", 
                                          original_indication == "THYMOMA AND THYMIC CARCINOMA" ~
                                            "THYM", 
                                          original_indication == "BREAST CANCER" ~
                                            "TNBC", # they are all TN
                                          str_detect(original_indication, 
                                                     "MELANOMA") ~
                                            "MELANOMA",
                                          str_detect(original_indication, 
                                                     "MESOTHELIOMA|PLEURAL") ~
                                            "MESOTHELIOMA",
                                          original_indication == "GASTRIC CANCER" ~
                                            "GC", 
                                          original_indication == "RENAL CELL CANCER" ~
                                            "RCC", 
                                          original_indication == "TESTICULAR CANCER" ~
                                            "TESTICULAR", 
                                          original_indication == "MALIGNANT FIBROUS HISTIOCYTOMA OF BONE" ~
                                            "MFHB", 
                                          original_indication == "FIBROLAMELAR HEPATOCARCINOMA" ~
                                            "HCC", 
                                          original_indication == "PANCREATIC CANCER" ~
                                            "PANCREATIC CA", 
                                          original_indication == "BLADDER CANCER" ~
                                            "BLADDER CA", 
                                          original_indication == "URETER CANCER" ~
                                            "URETER CA",
                                          original_indication == "PAPILLARY LUNG ADENOCARCINOMA" ~
                                            "LUNG ADENOCA",
                                          original_indication == "MERKEL CELL CARCINOMA" ~
                                            "MCC",
                                          original_indication %in% c("NEUROENDOCRINE TUMORS", 
                                                                     "GASTROINTESTINAL CARCINOID TUMOR") ~
                                            "NET", 
                                          TRUE ~ original_indication
                     
                   ))

PDL1_total_perc_immune_cells_stain <- derive_ZB(sdtm$zb, "CD274", "ICSTOT") %>% derive_pchgdata() %>% 
                                      mutate(VISIT = factor(VISIT, levels= c("Baseline", "On Treatment"), labels= c("Archival", "On Treatment"))) 

PDL1_total_perc_tum_inf_IC_in_tum_mass <- derive_ZB(sdtm$zb, "CD274", "TINFLICP") %>% derive_pchgdata() %>% 
                                      mutate(VISIT = factor(VISIT, levels= c("Baseline", "On Treatment"), labels= c("Archival", "On Treatment"))) 


PDL1_total_perc_ihc_pos_tum_cells <- derive_ZB(sdtm$zb, "CD274", "POSTUCP") %>% derive_pchgdata() %>% 
                                      mutate(VISIT = factor(VISIT, levels= c("Baseline", "On Treatment"), labels= c("Archival", "On Treatment"))) 
```

## CD8+ Absolute values

IRIS-IHC


```{r, echo=FALSE}

gg1 <- plot_ABS(CD8T, "", y_label = "CD8+ Density\n (cells/mm2)")
#fig6bleft

 # ggsave(filename = "Fig6B_left.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 12, unit= "cm", dpi = 300)

gg1 + ggsignif::geom_signif(comparisons = list(c("Baseline", "On Treatment")), map_signif_level = TRUE)

    

```

## CD8+ Single Agent: Patients with OT biopsies
```{r, echo=FALSE}
table_ZB(CD8T[CD8T$GROUP=='RO7122290 single agent',], "CD8+: Single Agent")
```

## CD8+ COMBO: Patients with OT biopsies
```{r, echo=FALSE}
table_ZB(CD8T[CD8T$GROUP=='RO7122290 + atezolizumab',], "CD8+: Combination")
```

## CD8+: Changes from baseline
```{r echo=FALSE, warning=FALSE}

gg2 <- plot_CHB(CD8T, "", y_label = "CD8+ \nFold change from baseline")


 # ggsave(filename = "Fig6B_right.pdf", path = "data/stm_2022_figures",
 #        width = 16, height = 12, unit= "cm", dpi = 300)

```

## Ki67+ CD8+ Absolute Values

```{r, echo=FALSE}


gg3 <- plot_ABS(KI67CD8T, "", y_label = "CD8+ Ki67+ Density\n (cells/mm2)")


 # ggsave(filename = "Fig6C_left.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 12, unit= "cm", dpi = 300)
 

```

## Ki67+ CD8+s: Single Agent patients with OT biopsies
```{r echo=FALSE, warning=FALSE}
table_ZB(KI67CD8T[KI67CD8T$GROUP=='RO7122290 single agent',], "Ki67+CD8+: Single Agent")
```

## Ki67+ CD8+: COMBO patients with OT biopsies
```{r echo=FALSE, warning=FALSE}
table_ZB(KI67CD8T[KI67CD8T$GROUP=='RO7122290 + atezolizumab',], "Ki67+CD8+: Combination")
```

## Ki67+ CD8+: Changes from baseline
```{r echo=FALSE, warning=FALSE}
gg4 <- plot_CHB(KI67CD8T,  "", y_label = "CD8+ Ki67+ \nFold change from baseline")


 # ggsave(filename = "Fig6C_right.pdf", path = "data/stm_2022_figures",
 #        width = 16, height = 12, unit= "cm", dpi = 300)
 

```




## FAP+ Absolute values
```{r, echo=FALSE}

gg5 <- plot_ABS(FAP_pos, "", "FAP+ \n(% Staining tumor area)")

#FigS6A

# ggsave(filename = "FigS6A.pdf", path = "data/stm_2022_figures",
#        width = 20, height = 12, unit= "cm", dpi = 300)

gg5 + ggsignif::geom_signif(comparisons = list(c("Baseline", "On Treatment")), map_signif_level = TRUE)


```

## FAP+ Single Agent: Patients with OT biopsies
```{r, echo=FALSE}
table_ZB(FAP_pos[FAP_pos$GROUP=='RO7122290 single agent',], "FAP+: Single Agent")
```

## CD8+ COMBO: Patients with OT biopsies
```{r, echo=FALSE}
table_ZB(FAP_pos[FAP_pos$GROUP=='RO7122290 + atezolizumab',], "FAP+: Combination")
```

## FAP+: Changes from baseline
```{r echo=FALSE, warning=FALSE}
gg6 <- plot_CHB(FAP_pos, "FAP+ Fold change", y_label = "FAP+ \nFold change from baseline")

```

## FAP+/panCK- Absolute values
```{r, echo=FALSE}

gg7 <- plot_ABS(FAP_ker_neg, "", "FAP+/panCK- \n(% Staining tumor area)")

# FigS6B

# ggsave(filename = "FigS6B.pdf", path = "data/stm_2022_figures",
#        width = 20, height = 12, unit= "cm", dpi = 300)


gg7 + ggsignif::geom_signif(comparisons = list(c("Baseline", "On Treatment")), map_signif_level = TRUE)


```

## FAP+/panCK- Single Agent: Patients with OT biopsies
```{r, echo=FALSE}
table_ZB(FAP_ker_neg[FAP_ker_neg$GROUP=='RO7122290 single agent',], "FAP+: Single Agent")
```

## CD8+/panCK- COMBO: Patients with OT biopsies
```{r, echo=FALSE}
table_ZB(FAP_ker_neg[FAP_ker_neg$GROUP=='RO7122290 + atezolizumab',], "FAP+: Combination")
```

## FAP+/panCK-: Changes from baseline
```{r echo=FALSE, warning=FALSE}
gg8 <- plot_CHB(FAP_ker_neg, "FAP+/panCk- Fold change", y_label = "FAP+/panCk- \nFold change from baseline")
 

```



## Risk ratio of increase in proliferating vs infiltrating CD8+ T cells

In 12 (63%) paired biopsies the densities of proliferating CD8+ T cells increased more than the densities of all infiltrating CD8 T cells. On average, the increase was 1.8-fold higher for proliferating CD8+ T cells than for all infiltrating CD8+ T cells (P=0.01, signed-rank test).

```{r echo=FALSE, warning=FALSE}
anl <- KI67CD8T %>% mutate(BASE.1 = BASE, OT.1 = AVAL) %>% filter(VISIT=='On Treatment' & is.na(RCHG)==FALSE) %>%
          select(USUBJID, ARMCD, EXDOSE, EXDOSU, AVALC, MHTERM, BASE.1, OT.1) %>%
        left_join(CD8T %>% mutate(BASE.2 = BASE, OT.2 = AVAL) %>% filter(VISIT=='On Treatment' & is.na(RCHG)==FALSE) %>%
          select(USUBJID, BASE.2, OT.2), 
          by = "USUBJID") %>%
        mutate(RR = (OT.1/BASE.1)/(OT.2/BASE.2))
summary(anl$RR)
table(anl$RR>1)
wilcox.test(anl$RR, mu = 1, alternative = "two.sided")
```

for fap

```{r echo=FALSE, warning=FALSE}
fapcks <- FAP_ker_neg %>% filter(VISIT=='On Treatment' & is.na(RCHG)==FALSE)
t.test(fapcks$BASE, fapcks$AVAL)

faps <- FAP_pos %>% filter(VISIT=='On Treatment' & is.na(RCHG)==FALSE)
t.test(faps$BASE, faps$AVAL)
tot_cd8_test <- CD8T %>% filter(VISIT=='On Treatment' & is.na(RCHG)==FALSE)
t.test(tot_cd8_test$BASE, tot_cd8_test$AVAL)
prl_cd8_test <- KI67CD8T %>% filter(VISIT=='On Treatment' & is.na(RCHG)==FALSE)
t.test(prl_cd8_test$BASE, prl_cd8_test$AVAL)
```
# Additional analysis





```{r plots, echo=FALSE}
#### Waterfall plot 
g_waterf_bestcha_pts <- function(anl, title){
  ##cut the USUBID down to last pt number
  anl$USUBJIDcut <- do.call(rbind, str_split(anl$USUBJID, "-"))[,length(str_split(anl$USUBJID, "-")[[1]])]
  ## keep only baseline per patient and use minPCHG
  anl <- anl %>% filter(ABLFL=='Y')
  
  
  groupcols <- c("Yes"= "royalblue", 
                 "No"= "red")
  
  N.CPI = c(as.character(dim(anl %>% filter(CPIFL=='Yes'))[1]),
            as.character(dim(anl %>% filter(CPIFL=='No' ))[1]))
  
  
  ##plot
  wat <- ggplot(data= anl, mapping = aes(x = reorder(USUBJIDcut, -minPCHG), y = minPCHG, fill=reorder(CPIFL, -minPCHG))) + 
    geom_bar(stat = "identity", position = "stack") +
    geom_hline(aes(yintercept = 20), linetype="dotted" , color = "black") +
    geom_hline(aes(yintercept = 0), linetype="dashed" , color = "black") +
    geom_hline(aes(yintercept = -30), linetype="dotted" , color = "black") +
    scale_fill_manual(name="CPI experience", values = groupcols) +
    scale_y_continuous(breaks=seq(-100, 100, 20), limits=c(-100,100)) +
    scale_x_discrete(labels = NULL) +
    xlab("") +
    ylab("Best change in target lesions from baseline (%)")  +
    ggtitle(title)  + 
    theme_classic() + 
    theme(axis.line.x=element_blank(),
          axis.text.x=element_blank(), 
          axis.ticks.x=element_blank()
  
          )
  return(wat)
}
#### Waterfall plot by best overall response
g_waterf_bestcha_pts_BOR <- function(anl, title){
  ##cut the USUBID down to last pt number
  anl$USUBJIDcut <- do.call(rbind, str_split(anl$USUBJID, "-"))[,length(str_split(anl$USUBJID, "-")[[1]])]
  ## keep only baseline per patient and use minPCHG
  anl <- anl %>% filter(ABLFL=='Y') %>% 
      mutate(BOR = fct_rev(fct_relevel(.f = BOR, levels = order_bor)))  
  
  

                
  N.BOR = c(as.character(dim(anl %>% filter(BOR=='CR'))[1]),
            as.character(dim(anl %>% filter(BOR=='PR' ))[1]),
            as.character(dim(anl %>% filter(BOR=='SD' ))[1]),
            as.character(dim(anl %>% filter(BOR=='PD' ))[1]),
            as.character(dim(anl %>% filter(BOR=='Missing' ))[1]))
   
   
 ##plot
 wat <- ggplot(data= anl, mapping = aes(x = reorder(USUBJIDcut, -minPCHG), y = minPCHG, fill=reorder(BOR, -minPCHG))) + 
    geom_bar(stat = "identity", position = "stack") +
    geom_hline(aes(yintercept = 20), linetype="dotted" , color = "black") +
    geom_hline(aes(yintercept = 0), linetype="dashed" , color = "black") +
    geom_hline(aes(yintercept = -30), linetype="dotted" , color = "black") +
    scale_fill_manual(name="Response by RECIST 1.1", values = respcolours, breaks = order_bor) +
    scale_y_continuous(breaks=seq(-100, 100, 20), limits=c(-100,100)) +
    scale_x_discrete(labels = NULL) +
    xlab("") +
    ylab("Best change in target lesions from baseline (%)")  +
    labs(caption = paste(N.BOR[1], "CR, ", 
                         N.BOR[2], "PR, ", 
                         N.BOR[3], "SD, ", 
                         N.BOR[4], "PD, and ", 
                         N.BOR[5], "patients with missing response included", "\n", 
                         "Data from", snapshot.date)) +
    ggtitle(title)  + 
    theme_classic() + 
    theme(axis.line.x=element_blank(),
          axis.text.x=element_blank(), 
          axis.ticks.x=element_blank()
  
          )
 
  return(wat)
}







plot_CHB_FC_ADY <- function(dat, title, palette = dosecols_viridis, names_facets = LBTEST, scales = "fixed", 
                            dodge_width = 0, alpha_points = 0.8){
 
  toplot <- dat %>% filter(ADY>=0 & ADY<=43)  %>%
              mutate(GROUP2 = factor(case_when(GROUP=="RO7122290 single agent" ~"1",
                                               GROUP=="RO7122290 + atezolizumab" & EXDOSE<500   ~"2",
                                               GROUP=="RO7122290 + atezolizumab" & EXDOSE>=500  ~"3"), 
                                     levels = c("1", "2", "3"),
                                     labels = c("RO7122290 single agent", 
                                                "RO7122290 + atezolizumab \n(45-260 mg)",
                                                "RO7122290 + atezolizumab \n(500-2000 mg)")))
  
  gg <- ggplot(toplot, aes(x=as.numeric(ADY), y=RCHG, group=USUBJID, colour=as.factor(EXDOSE))) +
        geom_boxplot(aes(x=as.numeric(ADY.1), group=ADY.1), colour="gray14", fill="gray85", outlier.shape=NA, size=0.4, alpha=0.5, width = 4) +
          geom_point(aes(fill=as.factor(EXDOSE)), size=1.5, alpha= alpha_points,  colour="gray14", position = position_dodge(width = dodge_width), shape = 21) +
          geom_line(size=0.5, alpha=0.5, position = position_dodge(width = dodge_width)) +          
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          geom_hline(aes(yintercept = 1), linetype="dashed" , color = "black") +
          geom_hline(aes(yintercept = 10), linetype="dotted" , color = "black") +
          geom_hline(aes(yintercept = 0.1), linetype="dotted" , color = "black") +
          scale_x_continuous(labels = c(1,8, 15, 22, 29, 36, 43), breaks = c(1,8, 15, 22, 29, 36, 43)) +
          scale_y_log10() +
          xlab("Days after first dose") +
          ylab("Fold change from baseline") +
          scale_colour_manual(name="RO7122290 dose (mg)", values = palette) +
          scale_fill_manual(name="RO7122290 dose (mg)", values = palette) +
          ggtitle(title) +
          theme_bw() +
          facet_grid(rows = vars({{names_facets}}), cols = vars(GROUP2), scales = scales) 
    print(gg)
  
}



boxplots_clin_paper <- function(input_data, 
                                     y_variable = AVAL,
                                     shape = GROUP,
                                     fill = AVALC, 
                                     alpha_fill = 0.8, 
                                     color_legend_title_string = "",
                                     shape_legend_title_string = "",
                                     y_axis_title_string = "", 
                                     palette_fill = respcolours, 	
                                     facet_variable_cols_as_string = NULL,
                                     facet_variable_rows_as_string = NULL 
                                     
) {
  
  
  
  data_of_interest <- input_data %>% 
                          filter(AVALC != "")

  
  respcolours <- respcolours[names(respcolours) %in% unique(input_data$AVALC)]
  
  
  group_shape <- group_shape[names(group_shape) %in% unique(input_data$GROUP)]

  p <- ggplot(data_of_interest, 
              aes(
                  x = fct_relevel(AVALC, rev(order_bor)),
                  y = {{y_variable}}
              )) +
    geom_boxplot(outlier.shape = NA, 
                 alpha = alpha_fill, 
                 aes(fill = {{fill}})
    ) +
    geom_jitter(height = 0, 
                width = 0.2,
                size = 3,
                alpha = 0.8,
                color = "black",
                fill = "#525252",
                aes(shape = {{shape}} 
                    )) +
    labs(
      x = "BOR", 
      y = paste(y_axis_title_string)
      
    ) +
    scale_fill_manual(values = palette_fill, name = "") +
    scale_shape_manual(values = group_shape, name = "") +

  
  
  if (is.null(facet_variable_cols_as_string)==FALSE) {
    
    
    p <- p + facet_wrap(c(facet_variable_cols_as_string, 
                          facet_variable_rows_as_string))
    
    
  }
  
  
  p + theme_bw()
  
}


dotplots_clin_paper <- function(input_data, 
                                     y_variable = AVAL,
                                     shape = GROUP,
                                     fill = AVALC, 
                                     alpha_fill = 0.8, 
                                     color_legend_title_string = "",
                                     shape_legend_title_string = "",
                                     y_axis_title_string = "", 
                                     palette_fill = respcolours, 	
                                     facet_variable_cols_as_string = NULL,
                                     facet_variable_rows_as_string = NULL 
                                     
) {
  
  
  
  data_of_interest <- input_data %>% 
                          filter(AVALC != "")
  

  
  respcolours <- respcolours[names(respcolours) %in% unique(input_data$AVALC)]
  
  
  group_shape <- group_shape[names(group_shape) %in% unique(input_data$GROUP)]
  
  data_of_interest$USUBJIDcut <- do.call(rbind, str_split(data_of_interest$USUBJID, "-"))[,length(str_split(data_of_interest$USUBJID, "-")[[1]])]

  p <- ggplot(data_of_interest, 
              aes(
                  x = reorder(USUBJIDcut, -{{y_variable}}),
                  y = {{y_variable}},
                  shape = {{shape}}, 
                    fill = {{fill}}
              )) +

    geom_point(
                size = 3,
                alpha = 0.8,
                color = "black"
                ) +
    labs(
      x = "", 
      y = paste(y_axis_title_string)
      
    ) +
    scale_fill_manual(values = respcolours, name = "") +
    scale_shape_manual(values = group_shape, name = "") +
    guides(fill = guide_legend(override.aes = list(shape = 21)))

  
  
  if (is.null(facet_variable_cols_as_string)==FALSE) {
    
    
    p <- p + facet_wrap(c(facet_variable_cols_as_string, 
                          facet_variable_rows_as_string), 
                        scales = "free_x")
    
    
  }
  
  
  p + theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.5))
  
}




```



## BP40087 Part A: Waterfall plot by CPI experience



```{r , echo = TRUE}
g_waterf_bestcha_pts(anl = SLD[SLD$GROUP=='RO7122290 single agent',], 
                     title = "RO7122290 single agent") +
    annotate("text", x = 53, y = -44, label = c("*")) 


# 
 # ggsave(filename = "Fig1A.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 14, unit= "cm", dpi = 300)
```

```{r , echo = TRUE}
g_waterf_bestcha_pts_BOR(anl = SLD[SLD$GROUP=='RO7122290 single agent',], 
                         title = "RO7122290 single agent") +
    annotate("text", x = 53, y = -44, label = c("*")) 
```


## BP40087 Part B/IMG: Waterfall plot by CPI experience

```{r , echo = TRUE}
g_waterf_bestcha_pts(anl = SLD[SLD$GROUP=='RO7122290 + atezolizumab',], 
                     title = "RO7122290 + atezolizumab")  +
    annotate("text", x = c(37,39), y = c(-43,-46), label = c("*", "*")) 

 # ggsave(filename = "Fig1B.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 14, unit= "cm", dpi = 300)
```




```{r , echo = TRUE}


g_waterf_bestcha_pts_BOR(anl = SLD[SLD$GROUP=='RO7122290 + atezolizumab',], 
                         title = "RO7122290 + atezolizumab") +
    annotate("text", x = c(37,39), y = c(-43,-46), label = c("*", "*")) 
                     
```


# Soluble 4-1-BB


Figure 3

```{r , echo = TRUE}
plot_CHB_FC_ADY(OTHBM %>% filter(LBTEST=="Soluble 4-1BB", ADY <40, VISITDY <40), "")

 # ggsave(filename = "Fig3.pdf", path = "data/stm_2022_figures",
 #        width = 24, height = 12, unit= "cm", dpi = 300)


```

# Cytokines Plot


```{r , echo = TRUE, fig.height= 14}


cytokines_abbreviations <- c("Interferon Gamma" = "IFN\u03B3", 
                             "Interleukin 6" = "IL-6", 
                             "Tumor Necrosis Factor" = "TNF\u03B1", 
                             "Soluble Interleukin 2 Receptor" = "sCD25")

CYT %>% filter(ADY.1 <40) %>% 
  mutate(cyt_short = str_replace_all(LBTEST, cytokines_abbreviations)) %>% 
  plot_CHB_FC_ADY("", names_facets = cyt_short)


CYT %>% filter(ADY.1 <40) %>% 
  mutate(cyt_short = str_replace_all(LBTEST, cytokines_abbreviations)) %>% 
  plot_CHB_FC_ADY("", names_facets = cyt_short, scales = "free_y", alpha_points = 0.4) 
# Fig5

 # ggsave(filename = "Fig5.pdf", path = "data/stm_2022_figures",
 #        width = 22, height = 20, unit= "cm", dpi = 300, device=cairo_pdf)
 
```



```{r, echo= TRUE}

CYT %>% filter(ADY.1 <40, 
               LBTEST == "Interferon Gamma") %>% 
      mutate(cyt_short = str_replace_all(LBTEST, cytokines_abbreviations)) %>%
  plot_CHB_FC_ADY("D", names_facets = cyt_short)


CYT %>% filter(ADY.1 <40, 
               LBTEST == "Interleukin 6") %>% 
      mutate(cyt_short = str_replace_all(LBTEST, cytokines_abbreviations)) %>%
  plot_CHB_FC_ADY("D", names_facets = cyt_short)

CYT %>% filter(ADY.1 <40, 
               LBTEST == "Soluble Interleukin 2 Receptor") %>% 
      mutate(cyt_short = str_replace_all(LBTEST, cytokines_abbreviations)) %>% 
  plot_CHB_FC_ADY("", names_facets = cyt_short)


CYT %>% filter(ADY.1 <40, 
               LBTEST == "Tumor Necrosis Factor") %>% 
    mutate(cyt_short = str_replace_all(LBTEST, cytokines_abbreviations)) %>% 
  plot_CHB_FC_ADY("", names_facets = cyt_short)

                 

```


# Biomarkers and responses

## FAP+/panCK-

```{r}




FAP_ker_neg %>% 
  filter(VISIT == "Baseline") %>% 
  boxplots_clin_paper(y_axis_title_string = "FAP+/panCk- (%)")

 # ggsave(filename = "Fig2C.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 14, unit= "cm", dpi = 300)


FAP_ker_neg %>% 
  filter(GROUP == "RO7122290 single agent") %>% 
  filter(VISIT == "Baseline") %>% 
  boxplots_clin_paper(y_axis_title_string = "FAP+/panCk- (%)")

FAP_ker_neg %>% 
  filter(GROUP == "RO7122290 + atezolizumab") %>% 
  filter(VISIT == "Baseline") %>% 
  boxplots_clin_paper(y_axis_title_string = "FAP+/panCk- (%)")


```


## FAP+

```{r}




FAP_pos %>%   
  filter(VISIT == "Baseline") %>% 
  boxplots_clin_paper(y_axis_title_string = "FAP+ (%)")

 # ggsave(filename = "FigS3A.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 12, unit= "cm", dpi = 300)

# figS3

FAP_pos %>% 
  filter(VISIT == "Baseline") %>% 
  filter(GROUP == "RO7122290 single agent") %>% 
  boxplots_clin_paper(y_axis_title_string = "FAP+ (%)")

FAP_pos %>% 
  filter(VISIT == "Baseline") %>% 
  filter(GROUP == "RO7122290 + atezolizumab") %>% 
  boxplots_clin_paper(y_axis_title_string = "FAP+ (%)")

```

## Total CD8

```{r}



CD8T %>% 
  filter(VISIT == "Baseline") %>% 
  boxplots_clin_paper(y_axis_title_string = "CD8+ (count/mm2)")
# FigS5A

 # ggsave(filename = "FigS5A.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 12, unit= "cm", dpi = 300)

CD8T %>% 
  filter(VISIT == "Baseline") %>% 
  filter(GROUP == "RO7122290 single agent") %>% 
  boxplots_clin_paper(y_axis_title_string = "CD8+ (count/mm2)")

CD8T %>% 
  filter(VISIT == "Baseline") %>% 
  filter(GROUP == "RO7122290 + atezolizumab") %>% 
  boxplots_clin_paper(y_axis_title_string = "CD8+ (count/mm2)")



```

## Proliferating CD8

```{r}




KI67CD8T %>% 
  filter(VISIT == "Baseline") %>% 
  boxplots_clin_paper(y_axis_title_string = "CD8+ Ki67+ (count/mm2)")

#FigS5b
 # ggsave(filename = "FigS5B.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 12, unit= "cm", dpi = 300)

KI67CD8T %>% 
  filter(VISIT == "Baseline") %>% 
  filter(GROUP == "RO7122290 single agent") %>% 
  boxplots_clin_paper(y_axis_title_string = "CD8+ Ki67+ (count/mm2)")

KI67CD8T %>% 
  filter(VISIT == "Baseline") %>% 
  filter(GROUP == "RO7122290 + atezolizumab") %>% 
  boxplots_clin_paper(y_axis_title_string = "CD8+ Ki67+ (count/mm2)")
      



                       
                       
```




## CD274 Absolute values


ICSTOT = Total Perc Immune Cells Stain <- Not used
POSTUCP = Perc IHC-Positive Tumor Cells <- call "PD-L1 SP263 (% tumor cells)"
TINFLICP = Perc Tum Infiltr Immune Cell In Tum Mass <- call "PD-L1 SP263 (% Tum infiltr IC in tumor)"


### Perc IHC-Positive Tumor Cells

 
```{r}



PDL1_total_perc_ihc_pos_tum_cells %>% 
    dotplots_clin_paper(y_axis_title_string  = "PD-L1 SP263 (% tumor cells)")


PDL1_total_perc_ihc_pos_tum_cells %>% 
    dotplots_clin_paper(y_axis_title_string  = "PD-L1 SP263 (% tumor cells)", facet_variable_cols_as_string = "GROUP")



PDL1_total_perc_ihc_pos_tum_cells %>% 
    filter(GROUP == "RO7122290 + atezolizumab") %>% 
    dotplots_clin_paper(y_axis_title_string  = "PD-L1 SP263 (% tumor cells)")
#Figs4btop

 # ggsave(filename = "FigS4B_top.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 10, unit= "cm", dpi = 300)

PDL1_total_perc_ihc_pos_tum_cells %>% 
    filter(GROUP == "RO7122290 single agent") %>% 
    dotplots_clin_paper(y_axis_title_string  = "PD-L1 SP263 (% tumor cells)")

#Figs4bbottom
 # ggsave(filename = "FigS4B_bottom.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 10, unit= "cm", dpi = 300)


PDL1_total_perc_ihc_pos_tum_cells %>% 
    boxplots_clin_paper(y_axis_title_string  = "PD-L1 SP263 (% tumor cells)")


PDL1_total_perc_ihc_pos_tum_cells %>% 
    filter(GROUP == "RO7122290 + atezolizumab") %>% 
    boxplots_clin_paper(y_axis_title_string  = "PD-L1 SP263 (% tumor cells)")









table_pdl1_tum_cells <- PDL1_total_perc_ihc_pos_tum_cells %>%
                        select(USUBJID,
                               GROUP,
                               VISIT,
                               ZBTARGET,
                               ZBTESTCD,
                               ZBTEST,
                               EXDOSE,
                               EXDOSU,
                               MHTERM,
                               result = AVAL.x,
                               PARAMU)
  
  
DT::datatable(data = table_pdl1_tum_cells, 
              filter = "top", 
              extensions = 'Buttons', 
              
              options = list(pageLength = 10, scrollX = TRUE, 
                             dom = 'Bfrtip', 
                   buttons = c('excel', "csv"))
              
              )
```



### Perc Tum Infiltr Immune Cell In Tum Mass



```{r}



PDL1_total_perc_tum_inf_IC_in_tum_mass %>% 
    dotplots_clin_paper(y_axis_title_string = "PD-L1 SP263 (% Tum infiltr IC in tumor)")

PDL1_total_perc_tum_inf_IC_in_tum_mass %>% 
    dotplots_clin_paper(y_axis_title_string = "PD-L1 SP263 (% Tum infiltr IC in tumor)", 
                        facet_variable_cols_as_string = "GROUP")


PDL1_total_perc_tum_inf_IC_in_tum_mass %>% 
    filter(GROUP == "RO7122290 + atezolizumab") %>% 
    dotplots_clin_paper(y_axis_title_string = "PD-L1 SP263 (% Tum infiltr IC in tumor)")
#FigS4Atop

# ggsave(filename = "FigS4A_top.pdf", path = "data/stm_2022_figures",
#        width = 20, height = 10, unit= "cm", dpi = 300)


PDL1_total_perc_tum_inf_IC_in_tum_mass %>% 
    filter(GROUP == "RO7122290 single agent") %>% 
    dotplots_clin_paper(y_axis_title_string = "PD-L1 SP263 (% Tum infiltr IC in tumor)")
#FigS4Abottom
# 
 # ggsave(filename = "FigS4A_bottom.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 10, unit= "cm", dpi = 300)

PDL1_total_perc_tum_inf_IC_in_tum_mass %>% 
    boxplots_clin_paper(y_axis_title_string = "PD-L1 SP263 (% Tum infiltr IC in tumor)")


PDL1_total_perc_tum_inf_IC_in_tum_mass %>% 
    filter(GROUP == "RO7122290 + atezolizumab") %>% 
    boxplots_clin_paper(y_axis_title_string = "PD-L1 SP263 (% Tum infiltr IC in tumor)")

table_pdl1_imm_cells_in_tum_mass <- PDL1_total_perc_tum_inf_IC_in_tum_mass %>% 
                        select(USUBJID, 
                               GROUP, 
                               VISIT, 
                               ZBTARGET, 
                               ZBTESTCD, 
                               ZBTEST, 
                               EXDOSE, 
                               EXDOSU, 
                               MHTERM, 
                               result = AVAL.x, 
                               PARAMU)
  
  
DT::datatable(data = table_pdl1_imm_cells_in_tum_mass, 
              filter = "top", 
              extensions = 'Buttons', 
              
              options = list(pageLength = 10, scrollX = TRUE, 
                             dom = 'Bfrtip', 
                   buttons = c('excel', "csv"))
              
              )


```



# Immunephenotype

```{r}






mi_p %>% 
  filter(!is.na(visit_name), 
         visit_name != "Archival") %>% 
  ggplot(
       aes(x = visit_name, 
           stratum = status, 
           alluvium = USUBJID, 
           fill = status, 
           label = status))+
  scale_fill_brewer(type = "qual", palette = "Set3") +
  geom_flow(stat = "alluvium", lode.guidance = "frontback",
            color = "darkgray") +
  scale_x_discrete(expand = c(.1, .1)) +
  geom_stratum(na.rm = TRUE) +
  geom_text(stat = "stratum", size = 3) +
  theme_classic()+
   theme(legend.position = "none") +
  labs(title = "Immunophenotype", 
       x = "", 
       y = "Number of participants")

complete_bsl_ot <- mi_p %>% 
                  filter(!is.na(visit_name), 
                          visit_name != "Archival") %>% 
                  group_by(USUBJID) %>% 
                  summarise(n = n()) %>% 
                  filter(n == 2) %>% 
                  select(USUBJID)


mi_p %>% 
  filter(!is.na(visit_name), 
         visit_name != "Archival", 
         USUBJID %in% complete_bsl_ot$USUBJID) %>% 
  ggplot(
       aes(x = visit_name, 
           stratum = status, 
           alluvium = USUBJID, 
           fill = GROUP, 
           label = status))+
  scale_fill_manual(values = cols_allu[c(6:8,10:11)]) +
  geom_flow(stat = "alluvium",
            lode.guidance = "frontback",
            color = "darkgray", alpha = 1) +
  scale_x_discrete(expand = c(.1, .1)) +
  geom_stratum(na.rm = TRUE, 
               aes(fill = status)) +
  geom_text(stat = "stratum", size = 3) +
  theme_classic()+
   theme(legend.title = element_blank()) +
  labs(title = "", 
       x = "", 
       y = "Number of participants")

#figs7

# ggsave(filename = "FigS7.pdf", path = "data/stm_2022_figures",
#        width = 20, height = 14, unit= "cm", dpi = 300)

mi_p %>% 
  filter(!is.na(visit_name), 
         visit_name != "Archival", 
         USUBJID %in% complete_bsl_ot$USUBJID) %>% 
  ggplot(
       aes(x = visit_name, 
           stratum = status, 
           alluvium = USUBJID, 
           fill = BOR, 
           label = status))+
  scale_fill_manual(values = cols_allu) +
  geom_flow(stat = "alluvium",
            lode.guidance = "frontback",
            color = "darkgray", alpha = 1) +
  scale_x_discrete(expand = c(.1, .1)) +
  geom_stratum(na.rm = TRUE, 
               aes(fill = status)) +
  geom_text(stat = "stratum", size = 3) +
  theme_classic()+
   theme(legend.title = element_blank()) +
  labs(title = "Immunophenotype", 
       x = "", 
       y = "Number of participants")


mi_p %>% 
  filter(!is.na(visit_name), 
         visit_name != "Archival") %>% 
  ggplot(
       aes(x = visit_name, 
           stratum = status, 
           alluvium = USUBJID, 
           fill = GROUP, 
           label = status))+
  scale_fill_manual(values = cols_allu) +
  geom_flow(stat = "alluvium",
            lode.guidance = "frontback",
            color = "darkgray", alpha = 1) +
  scale_x_discrete(expand = c(.1, .1)) +
  geom_stratum(na.rm = TRUE, 
               aes(fill = status)) +
  geom_text(stat = "stratum", size = 3) +
  theme_classic()+
   theme(legend.title = element_blank()) +
  labs(title = "Immunophenotype", 
       x = "", 
       y = "Number of participants")


mi_p %>% 
  filter(!is.na(visit_name), 
         visit_name != "Archival") %>% 
  ggplot(
       aes(x = visit_name, 
           stratum = status, 
           alluvium = USUBJID, 
           fill = BOR, 
           label = status))+
  scale_fill_manual(values = cols_allu) +
  geom_flow(stat = "alluvium",
            lode.guidance = "frontback",
            color = "darkgray", alpha = 1) +
  scale_x_discrete(expand = c(.1, .1)) +
  geom_stratum(na.rm = TRUE, 
               aes(fill = status)) +
  geom_text(stat = "stratum", size = 3) +
  theme_classic()+
   theme(legend.title = element_blank()) +
  labs(title = "Immunophenotype", 
       x = "", 
       y = "Number of participants")


```




```{r old.functions, echo = TRUE}
plot_CHB_med <- function(dat, title){
    dat$ADY[dat$VISIT=='Baseline'] = 0
    toplot <- dat %>% filter(ADY>=0)  %>%
              mutate(LBTEST = factor(case_when(LBTEST=="Soluble Interleukin 2 Receptor" ~"1",
                                                LBTEST=="Tumor Necrosis Factor" ~"2",
                                                LBTEST=="Interferon Gamma" ~"3",
                                                LBTEST=="Interleukin 6" ~"4"),
                                     levels = c("1", "2", "3", "4"),
                                     labels = c("Soluble CD25",
                                                "Tumor Necrosis Factor",
                                                "Interferon Gamma",
                                                "Interleukin 6"))) %>%
              filter(is.na(RCHG)==FALSE) %>% group_by(LBTEST, GROUP, EXDOSE, ADY) %>%
              summarize(n1=n(),
                        med1=median(RCHG, na.rm=TRUE),
                        iqrl=quantile(RCHG, probs=0.25,na.rm=TRUE),
                        iqrh=quantile(RCHG, probs=0.75,na.rm=TRUE)) %>%
              ungroup() %>%
              mutate(GROUP2 = factor(case_when(GROUP=="RO7122290 single agent" ~"1",
                                               GROUP=="RO7122290 + atezolizumab" & EXDOSE<500   ~"2",
                                               GROUP=="RO7122290 + atezolizumab" & EXDOSE>=500  ~"3"),
                                     levels = c("1", "2", "3"),
                                     labels = c("RO7122290 single agent",
                                                "RO7122290 + atezolizumab \n(45-260 mg)",
                                                "RO7122290 + atezolizumab \n(500-2000 mg)")))
    gg <- ggplot(toplot, aes(x=as.numeric(ADY), y=med1, group=as.factor(EXDOSE), colour=as.factor(EXDOSE))) +
          geom_line(size=1, alpha=0.8, position = position_dodge(width = 0.3)) +
          geom_errorbar(aes(ymin=iqrl, ymax=iqrh), width=.1 , position = position_dodge(width = 0.3)) +
          geom_point(size=1.5, alpha=0.9, position = position_dodge(width = 0.3)) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          geom_hline(aes(yintercept = 1), linetype="dashed" , color = "black") +
          geom_hline(aes(yintercept = 10), linetype="dotted" , color = "black") +
          geom_hline(aes(yintercept = 0.1), linetype="dotted" , color = "black") +
          #geom_text(aes(label=n1), size = 3 ,vjust=0.5, hjust = 0.5, color= "black", position = position_dodge(width = 0.3)) +
          scale_x_continuous(labels = c(1,8, 15, 22, 29, 36, 43), breaks = c(1,8, 15, 22, 29, 36, 43)) +
          scale_y_log10() +
          xlab("Days after first dose") +
          ylab("Fold change from baseline") +
          scale_colour_manual(name="RO7122290 dose (mg)", values = dosecols) +
          ggtitle(title) +
          theme_bw() +
          facet_grid(rows = vars(LBTEST), cols = vars(GROUP2)) +
          labs(caption = paste("Plotted are median fold changes over time,", "\n",
                               "vertical lines denote interquartile ranges.", "\n",
                               "Data from", snapshot.date))
    print(gg)
}
```

## Other relations

```{r}


wide_others <- FAP_ker_neg %>% 
                bind_rows(KI67CD8T) %>% 
                bind_rows(CD8T) %>% 
                select(USUBJID, 
                       ARMCD, 
                       VISIT, 
                       ZBTARGET,
                       ZBORRES, 
                       EXDOSE, 
                       AVALC, 
                       MHTERM, 
                       BASE, 
                       AVAL, 
                       CHG) %>% 
                pivot_wider(names_from = c(VISIT, 
                       ZBTARGET),
                            values_from = c(
                       ZBORRES,BASE, 
                       AVAL, 
                       CHG)
                            )



scatterplot_wide <- function( data,
                             x_variables, 
                             y_variables,
                             palette_fill = respcolours[1:4], 
                             spearman_corr = FALSE,
                             position_spearman_legend_y = "top",
                             position_spearman_legend_x = "left",
                             shape = NULL,
                             fill = NULL,
                             y_label = "ADD LABEL", 
                             x_label = "ADD LABEL",
                             shape_title_string = "",
                             alpha = 1) {
  
  
  
  g <- ggplot(data = data, aes(x = {{x_variables}}, 
                               y = {{y_variables}} )) +
    
    geom_point(aes(
      #shape = {{shape}},
      fill = {{fill}}
    ),
    size = 2,
    shape = 21,
    alpha = alpha) +
    
    scale_fill_manual(values = palette_fill, 
                      name = "", drop = TRUE) +
    # scale_shape_manual(values=c(21, 24, 22, 23,4), 
    #                    name = "") +
    theme(legend.position = "right") +
    labs(y = y_label, 
         x = x_label
           ) +
    theme_bw()
  
  
  if (spearman_corr == TRUE) {
    g <- g  + stat_cor(method="spearman", 
                       label.x.npc = position_spearman_legend_x, 
                       label.y.npc = position_spearman_legend_y,
                       aes(label = paste(sub("R = ","Spearman=",..r.label..),
                                         "\n",..p.label..)),
                       output.type="text"
    )
    
  }
  
g

  
}



ggplot(wide_others, 
       aes(x = `CHG_On Treatment_MKI67+ CD8A+`, 
           y = `BASE_Baseline_FAP+/KRT-`)) +
  geom_point(aes(fill = AVALC), shape = 21, size = 2) +
  scale_fill_manual(name="Response by RECIST 1.1", values = respcolours, breaks = order_bor) +
  theme_bw()



scatterplot_wide(wide_others, 
                 x_variables =  `CHG_On Treatment_MKI67+ CD8A+`, 
                 y_variables = `BASE_Baseline_FAP+/KRT-`, 
                 fill = AVALC, 
                 y_label = "Baseline FAP+/panCK-", 
                 x_label = "Ki67+CD8+ Change from baseline", 
                 spearman_corr = TRUE)


scatterplot_wide(wide_others, 
                 x_variables =  `CHG_On Treatment_(MKI67+ CD8A+)+(MKI67- CD8A+)`, 
                 y_variables = `BASE_Baseline_FAP+/KRT-`, 
                 fill = AVALC, 
                 y_label = "Baseline FAP+/panCK-", 
                 x_label = "Total CD8+ Change from baseline", 
                 spearman_corr = TRUE)
# Figs3c

 # ggsave(filename = "FigS3C.pdf", path = "data/stm_2022_figures",
 #        width = 16, height = 12, unit= "cm", dpi = 300)


scatterplot_wide(wide_others, 
                 x_variables =  `CHG_On Treatment_MKI67+ CD8A+`, 
                 y_variables = `BASE_Baseline_(MKI67+ CD8A+)+(MKI67- CD8A+)`, 
                 fill = AVALC, 
                 y_label = "Baseline Total CD8+", 
                 x_label = "Ki67+CD8+ Change from baseline", 
                 spearman_corr = TRUE)

scatterplot_wide(wide_others, 
                 x_variables =  `CHG_On Treatment_(MKI67+ CD8A+)+(MKI67- CD8A+)`, 
                 y_variables = `BASE_Baseline_(MKI67+ CD8A+)+(MKI67- CD8A+)`, 
                 fill = AVALC, 
                 y_label = "Baseline Total CD8+", 
                 x_label = "Total CD8+ Change from baseline", 
                 spearman_corr = TRUE)

# FigS5C
 # ggsave(filename = "FigS5C.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 12, unit= "cm", dpi = 300)
```


## FAP at baseline by indication

Figure 2B

```{r}

FAP_ker_neg %>% 
  filter(VISIT == "Baseline", 
         AVALC != "") %>% 
  ggplot(aes(y = reorder(indication, AVAL), 
               x = AVAL)) +
    geom_boxplot(outlier.shape = NA, 
                 alpha = 0.8, 
                 fill = "#d9d9d9"
    ) +
    geom_jitter(alpha = 0.8, 
                     size = 2,
                shape = 21,
                height = 0.2,
                     aes(
                         fill = AVALC)
    ) +
    labs(#x = paste(as_label(enquo(x_variable)))
      x = paste("FAP+/panCk- (%)"), 
      y = ""
      
    ) +
    scale_fill_manual(values = respcolours, 
                      name = "") + theme_bw()



 # ggsave(filename = "Fig2B.pdf", path = "data/stm_2022_figures",
 #        width = 20, height = 12, unit= "cm", dpi = 300)




```





# TTE analysis


```{r}



il8_bsl <- OTHBM %>% 
              filter(PARAMCD == "INTLP8SI", 
                     AVISIT == "BASELINE") %>% 
  select(USUBJID, 
         il8_bsl = AVAL)

fap_krt_bsl <- FAP_ker_neg %>% 
                filter(VISIT == "Baseline") %>% 
  select(USUBJID, 
         fap_krt = ZBORRES)

total_cd8_bsl <- CD8T %>% 
                filter(VISIT == "Baseline") %>% 
  select(USUBJID, 
         cd8 = ZBORRES)

cpi_status_all <- adsl %>%
              select(USUBJID,
                     CPIFL) %>% 
              distinct() %>% 
              mutate(`CPI Experience` = case_when(CPIFL == "Y" ~ "Yes", 
                               CPIFL == "N" ~ "No",
                               TRUE ~ "Unknown (PROB ERROR)")) %>% 
              select(-CPIFL)


neut_lymph_bsl <- OTHBM %>% 
              filter(PARAMCD %in% c("NEUTRSI", "LYMPHSI"), 
                     AVISIT == "BASELINE", 
                     LBTPT == "PREDOSE") %>%
  select(USUBJID, 
         PARAMCD,
         AVAL) %>% 
  pivot_wider(id_cols = USUBJID, 
              names_from = PARAMCD, 
              values_from = AVAL) %>% 
  mutate(neut_over_lym = NEUTRSI/LYMPHSI) %>% 
  select(USUBJID, 
         neut_over_lym)

# read external data in separate script

data_cox <- entimice_tte %>% 
    left_join(il8_bsl, 
              by = "USUBJID") %>% 
  left_join(total_cd8_bsl, 
              by = "USUBJID") %>% 
  left_join(fap_krt_bsl, 
              by = "USUBJID")  %>% 
  left_join(cpi_status_all, 
              by = "USUBJID")%>% 
  left_join(external_info_clin_sci, 
              by = "USUBJID")  %>% 
  left_join(neut_lymph_bsl, 
              by = "USUBJID")  %>% 
  mutate(`IL8 level`  = case_when(il8_bsl >= 20 ~ "High", 
                                                                             il8_bsl < 20 ~ "Low", 
                                                                             is.na(il8_bsl) ~ "Unknown"),
         `CD8 over 350 bsl` = case_when(cd8 > 350 ~ "Yes", 
                                                   cd8 <=  350 ~ "No", 
                                                   TRUE ~ "Unknown"),
         `FAP+/panCk- level` = case_when(fap_krt >= 20 ~ "High", 
                                                                             fap_krt < 20 ~ "Low", 
                                                                             is.na(fap_krt) ~ "Unknown"),
         `Neutrophils/lymphocytes` = case_when(neut_over_lym >= 4 ~ "High", 
                                                                             neut_over_lym < 4 ~ "Low", 
                                                                             is.na(neut_over_lym) ~ "Unknown")
         ) %>% 
        mutate(`IL8 level` = fct_relevel(`IL8 level`, c("Low", "High", "Unknown")),
                         `FAP+/panCk- level` = fct_relevel( `FAP+/panCk- level`, c("Low", "High", "Unknown")), 
                         `Neutrophils/lymphocytes` = fct_relevel( `Neutrophils/lymphocytes`, c("Low", "High", "Unknown")))


res.cox.entimice <- coxph(Surv(AVAL, censor_reverse) ~ `IL8 level` +
                            `FAP+/panCk- level` +
                            `Liver metastasis` +
                               `CD8 over 350 bsl` +
                            `CPI Experience` +
                             `LOT <=4` +
                            `Neutrophils/lymphocytes`, 
                          data = data_cox
                )

summary(res.cox.entimice)
ggforest(res.cox.entimice, 
         fontsize = 1)

# ggsave(here::here("data/stm_2022_figures/Fig8.pdf"),
#        width = 30, height = 20, unit= "cm", dpi = 300)
```


